---
title: "sword_shield_pokemon"
author: "Caspar Hu"
date: "1/16/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(rvest)
library(dplyr)
library(tidyr)
library(stringr)
library(ggplot2)
```

## Web Scraping 
```{r}
link <- "https://www.serebii.net/swordshield/galarpokedex.shtml"
pokemon_ss <- read_html(link) %>%  # web scraping
  html_nodes("table.tab")%>%
  html_table(fill=TRUE)%>%.[[2]]
```

```{r}
link_2 <- "https://www.ign.com/wikis/pokemon-sword-shield/List_of_Pokemon_(Pokedex)#Galarian_Pokedex"
pokemon_type <- read_html(link_2) %>%  # web scraping
  html_nodes("table")%>%   
  html_table(fill=TRUE)%>%.[[1]]
```

## Data Cleaning
```{r}
colnames(pokemon_ss)

head(pokemon_ss)

pokemon_ss <- pokemon_ss[,-6]   # remove sixth column
pokemon_ss <- pokemon_ss[,-2:-3] # remove second, third column

colnames(pokemon_ss) <- c("No.","Name","Abilities","HP","Attack","Defence","Sp_Attack","Sp.Defence","Speed")
# change column name

pokemon_ss <- pokemon_ss[-1:-2,] # remove first, second rows

pokemon_ss <- na.omit(pokemon_ss) # remove all the NA rows 
```

```{r}
pokemon_ss$No. <-  str_replace_all(pokemon_ss$No.,"\\#","") # remove # at the beginning of the No. column

pokemon_ss$No. <- as.numeric(pokemon_ss$No.) # cahnge into numeric
```
```{r}
unique(pokemon_ss$No.)   # there are 400 unique No.
unique(pokemon_type$`#`) # there are 400 unique observations

pokemon_list <- merge(pokemon_type, pokemon_ss, by.x = "#", by.y = "No.")

colnames(pokemon_list) <- tolower(colnames(pokemon_list))

pokemon_list <- pokemon_list%>%
  rename(No.="#")
  
pokemon_list <- pokemon_list%>%  
  rename(locations_found="location(s) found")# rename the column name

unique(pokemon_list$No.) # there are 400 unique observations

summary(as.factor(pokemon_list$No.)) # use factor we can see which # has duplicates No.182 and No.327

# we can pull No.182 and No.327 out 

pokemon_list[pokemon_list$No.==182,]
pokemon_list[pokemon_list$No.==327,]

#Both pokemons have different types although they do have the same stats, we still decide to keep them in the data set
```

```{r}
summary(pokemon_list)
head(pokemon_list)

colnames(pokemon_list) <- c("No.","pokemon_name","type","locations_found","name",
                             "abilities","hp","attack", "defence","sp_attack",
                             "sp_defence","speed")


pokemon_list <- pokemon_list[,-5] # remove duplicate column

col_order <- c("No.","pokemon_name","type",
               "abilities","locations_found","hp","attack", 
               "defence","sp_attack",
               "sp_defence","speed")

pokemon_list <- pokemon_list[, col_order]

pokemon_list <- pokemon_list%>%
  mutate(No.=as.character(No.),
         type=as.factor(type),
         abilities=as.factor(abilities),
         hp=as.numeric(hp),
         attack=as.numeric(attack),
         defence=as.numeric(defence),
         sp_attack=as.numeric(sp_attack),
         sp_defence=as.numeric(sp_defence),
         speed=as.numeric(speed))

summary(pokemon_list)
```

```{r}
pokemon_list$type <- str_replace_all(pokemon_list$type,
                                     "(?<=[a-z])(?=[A-Z])", #find small letters connected with capital letter                                                             
                                     "\n") # separate the pattern with space
pokemon_list <-  pokemon_list%>%
  separate(col=type,
           into=c("type_1","type_2"), # separate the type column into two columns between space 
           sep="\n")

pokemon_list <-  pokemon_list%>%  
mutate(type_1=as.factor(type_1),
         type_2=as.factor(type_2))

summary(pokemon_list)

print(unique(pokemon_list$type_1))  # we found a typo there is one category named poison and the other named posion
print(unique(pokemon_list$type_2)) # no duplicates in this column, but NA does not help us too much let's change those NAs into None

pokemon_list$type_1[pokemon_list$type_1=="Posion"] <- "Poison" # change type_1 "posion" into correct name 

pokemon_list <- pokemon_list%>%
  mutate(type_2=as.character(type_2))%>%
  mutate(type_2=if_else(is.na(type_2),"None",type_2))%>% # change NAs to None
  mutate(type_2=as.factor(type_2))
```

```{r}
pokemon_list_long <- pokemon_list%>%  
  pivot_longer(cols=hp:speed,   # transform into long table
               names_to="stats_type",
               values_to="score")
```


## Data Analysis 

Personally I love water type Pokemon, so we are going to filter all the water type pokemon first.

```{r}
water <- pokemon_list%>%
  filter(type_1=="Water" | type_2=="Water")

summary(water)



```



```{r}
ggplot(water, aes(x = attack , y = defence , label = pokemon_name)) +
  ggrepel::geom_text_repel(size = 2) +
  theme_minimal()
```
```{r,fig.width=8}
ggplot(pokemon_list_long, aes(x = score , y = stats_type)) +
  geom_boxplot() + 
  theme_minimal()+
  facet_wrap(~type_1)
```
```{r}
library(viridis)
library(hrbrthemes)
# Plot
pokemon_list_long %>%
  ggplot( aes(x=score, y=stats_type,fill=stats_type)) +
    geom_boxplot() +
    scale_fill_viridis(discrete = TRUE, alpha=0.8, option="A") +
    geom_jitter(color="black", size=0.4, alpha=0.) +
    theme_ipsum() +
    theme(
      legend.position="none",
      plot.title = element_text(size=12)
    ) +
    ggtitle("All Pokemon in Galar Region Stats Boxplot") +
    xlab("Stats Score")+
    ylab("")
```
```{r}
# Plot
pokemon_list_long %>%
  ggplot( aes(x=score, y=stats_type,fill=stats_type)) +
    geom_violin(width=1) +
    geom_boxplot(width=0.1, color="grey", alpha=0.2) +
    scale_fill_viridis(discrete = TRUE) +
    theme_ipsum() +
    theme(
      legend.position="none",
      plot.title = element_text(size=12)
    ) +
    ggtitle("All Pokemon in Galar Region Stats Violin Boxplot (Wrapped)") +
    xlab("Stats Score")+
    ylab("")
```

```{r}
primary_type <- pokemon_list%>%
  group_by(type_1)%>%
  summarise(count=n())%>%
  arrange(-count)

g1 <- ggplot(data=primary_type, aes(x=type_1,y=count,fill=type_1))+
  geom_bar(position="dodge",stat="identity") +
  coord_flip()+
  ggtitle("Primary Type in Galar Region")+
  xlab("Primary Type")+
  ylab("Number of Pokemons")
  

#sort the bargraph
ggplot(data=primary_type, aes(x=reorder(type_1,-count),
                              y=count,fill=type_1))+
  geom_bar(position="dodge",stat="identity") +
  coord_flip()+
  ggtitle("Primary Type in Galar Region")

secondary_type <- pokemon_list%>%
  group_by(type_2)%>%
  summarise(count=n())%>%
  arrange(-count)
  
g2 <- ggplot(data=secondary_type, aes(x=type_2,y=count,fill=type_2))+
  geom_bar(position="dodge",stat="identity") +
  coord_flip()+
  ggtitle("Secondary Type in Galar Region")+
  xlab("Secondary Type")+
  ylab("Number of Pokemons")
  
#sort the bargraph
ggplot(data=secondary_type, aes(x=reorder(type_2,-count),
                              y=count,fill=type_2))+
  geom_bar(position="dodge",stat="identity") +
  coord_flip()+
  ggtitle("Secondary Type in Galar Region")

g1 # Primary Type Count
g2 # Secondary Type Count

```

```{r}
library(corrplot)
pokemon_list %>%
  select(where(is.numeric))%>%  # find all the numeric columns 
  cor()%>%
  corrplot(method = "number")
```
```{r}
library(GGally)
library(dplyr)
pokemon_stats <- pokemon_list%>%
  select(where(is.numeric))

ggpairs(pokemon_stats) 
```

```{r}
ggplot(pokemon_list, aes(x=attack, y=sp_attack)) + 
    geom_point(size=2,color="blue")
```
Heat map for all water type pokemon including primary and secondary type as water
```{r}
# Libraries
library(tidyverse)
library(hrbrthemes)
library(viridis)
library(plotly)

# Matrix format
mat <- water
rownames(mat) <- mat[,2]
mat <- mat[,7:12]
mat <- as.matrix(mat)

library(heatmaply)
heatmap(mat, Colv = NA, Rowv = NA, scale="column")

p <- heatmaply(mat, 
        dendrogram = "none",
        xlab = "", ylab = "", 
        main = "",
        scale = "column",
        margins = c(5,5),
        grid_color = "white",
        grid_width = 1,
        titleX = FALSE,
        hide_colorbar = TRUE,
        branches_lwd = 1,
        label_names = c("Country", "Feature:", "Value"),
        fontsize_row = 5, fontsize_col = 5,
        labCol = colnames(mat),
        labRow = rownames(mat),
        heatmap_layers = theme(axis.line=element_blank())
        )
p
```

## K means cluster 

```{r}
library(tidyr)
library(cluster)
library(factoextra) # clustering algorithms & visualization
library(sparcl) # Sparse Clustering
```

```{r Scale Data}
pokemon_new <- pokemon_list[,c("pokemon_name","hp","attack","defence","sp_attack","sp_defence","speed")]
# scale stats data 
pokemon_scale <- scale(pokemon_new[,2:7])

# Add pokemon names back to data frame
pokemon_new <- cbind.data.frame(pokemon_new$pokemon_name, pokemon_scale)  

names(pokemon_new)[1] <- "pokemon_name"  # Fix name of team column
```

### Initial Clustering

```{r fit kmeans}
set.seed(111111) # Set seed for reproducibility
fit_1 <- kmeans(x = pokemon_new[,2:7], # Set data as explantory variables 
                centers = 4,  # Set number of clusters
                nstart = 25, # Set number of starts
                iter.max = 100 ) # Set maximum number of iterations to use
```

```{r Extract results}
# Extract clusters
clusters_1 <- fit_1$cluster

# Extract centers
centers_1 <- fit_1$centers

```

Lets first check how many samples in each cluster:

```{r Check cluster numbers}
# Check samples per cluster
summary(as.factor(clusters_1))
```

Here we see that we have 83 samples in cluster 1, 92 in cluster 2, 64 in cluster 3, and 163 in cluster 4. We can view the pokemons in each cluster as follows:

```{r Check pokemons per cluster}
# Check pokemons in cluster 1
cat("Cluster 1 Pokemon:\n")
pokemon_new$pokemon_name[clusters_1 == 1]
# Check teams in cluster 2
cat("Cluster 2 Pokemon:\n")
pokemon_new$pokemon_name[clusters_1 == 2]
# Check teams in cluster 3
cat("Cluster 3 Pokemon:\n")
pokemon_new$pokemon_name[clusters_1 == 3]
# Check teams in cluster 4
cat("Cluster 4 Pokemon:\n")
pokemon_new$pokemon_name[clusters_1 == 4]
```
```{r}
# Create vector of clusters
cluster <- c(1: 4)
# Extract centers
center_df <- data.frame(cluster, centers_1)

# Reshape the data

center_reshape <- gather(center_df, # set the data 
                         stats, # the col name for the select columns' name
                         values, # the col name for the select columns' value
                         hp:speed) # select columns 

g_heat_1 <- ggplot(data = center_reshape, aes(x = stats, y = cluster, fill = values)) +
  scale_y_continuous(breaks = seq(1, 4, by = 1)) +
  geom_tile() +
  coord_equal() + 
  theme_set(theme_bw(base_size = 22) ) +
  scale_fill_gradient2(low = "blue", # Choose low color
                       mid = "white", # Choose mid color
                       high = "red", # Choose high color
                       midpoint =0, # Choose mid point
                       space = "Lab", 
                       na.value ="grey", # Choose NA value
                       guide = "colourbar", # Set color bar
                       aesthetics = "fill") + # Select aesthetics to apply
  coord_flip()
g_heat_1

```
Overall, cluster 1 performs much more better than other three clusters. Pokemons of cluster 1 are generally faster compared to pokemons in other three clusters. Pokemons of cluster 2 are better at physical attack and has a higher health value. Pokemons of cluster 3 are good at dealing with special attack due to higher special defense. And cluster 4 performs generally worse at all six stats.

### Method 1 : Calculate Optimal Cluster Number (Elbow Method)

Let's try to calculate the optimal cluster number to see if we have a better set of clusters.   
In general, while we increase the number of clusters, error will always decrease. The optimal cluster number is the one where rate of improvement in performance begins to decrease. And that is how we can find the optimal cluster number.

```{r Check Cluster Number}
# Create function to try different cluster numbers
kmean_withinss <- function(k) {
  cluster <- kmeans( x = pokemon_new[,2:7],  # Set data to use
                    centers = k,  # Set number of clusters as k, changes with input into function
                    nstart = 25, # Set number of starts
                    iter.max = 100) # Set max number of iterations
  return (cluster$tot.withinss) # Return cluster error/within cluster sum of squares
}

# Set maximum cluster number
max_k <- 20
# Run algorithm over a range of cluster numbers
wss <- sapply(2:max_k, kmean_withinss)

# Create a data frame to plot the graph
elbow <-data.frame(2:max_k, wss)

# Plot the graph with gglop
g_1 <- ggplot(elbow, aes(x = X2.max_k, y = wss)) +  
  theme_set(theme_bw(base_size = 22) ) +
  geom_point(color = "blue") +
  geom_line() +
  scale_x_continuous(breaks = seq(1, 20, by = 1)) +
  labs(x = "Number of Clusters", y="Within Cluster \nSum of Squares") +
  theme(panel.grid.major = element_blank(), # Turn of the background grid
        panel.grid.minor = element_blank(),
        panel.border = element_blank(),
        panel.background = element_blank()) 
g_1

```
It seems that the optimal number of clusters is five. So, let's try it out

```{r fit kmeans 2}
set.seed(111111) # Set seed for reproducibility
fit_2 <- kmeans(x = pokemon_new[,2:7], # Set data as explantory variables 
                centers = 5,  # Set number of clusters
                nstart = 25, # Set number of starts
                iter.max = 100 ) # Set maximum number of iterations to use
```

```{r Extract results 2}
# Extract clusters
clusters_2 <- fit_2$cluster

# Extract centers
centers_2 <- fit_2$centers

# Check samples per cluster
summary(as.factor(clusters_2))

```

```{r Check pokemons per cluster}
# Check pokemons in cluster 1
cat("Cluster 1 Pokemon:\n")
pokemon_new$pokemon_name[clusters_2 == 1]
# Check pokemons in cluster 2
cat("Cluster 2 Pokemon:\n")
pokemon_new$pokemon_name[clusters_2 == 2]
# Check pokemons in cluster 3
cat("Cluster 3 Pokemon:\n")
pokemon_new$pokemon_name[clusters_2 == 3]
# Check pokemons in cluster 4
cat("Cluster 4 Pokemon:\n")
pokemon_new$pokemon_name[clusters_2 == 4]
# Check pokemons in cluster 4
cat("Cluster 5 Pokemon:\n")
pokemon_new$pokemon_name[clusters_2 == 5]
```

Lets check the quality of our clustering solution using a silhouette plot:

```{r Quality Check, fig.width=8}
dis = dist(pokemon_new[2:7])^2
#op <- par(mfrow= c(1,1), oma= c(0,0, 3, 0),
          #mgp= c(1.6,.8,0), mar= .1+c(4,2,2,2))   
sil = silhouette (fit_1$cluster , dis, full = TRUE)
plot(sil)
sil = silhouette (fit_2$cluster , dis, full = TRUE)
plot(sil)

# silhouette coefficients (as these values are referred to as) near +1 indicate that the sample is far away from the neighboring clusters. A value of 0 indicates that the sample is on or very close to the decision boundary between two neighboring clusters and negative values indicate that those samples might have been assigned to the wrong cluster.
```

From four clusters' silhouette plot, we can see both cluster 2 and 3 have samples that do not belong to their groups, while it is even worse in five clusters' silhouette plot, Clusters 1, 2 and 5 have samples that do not belong to their groups. 

### Checking Clustering Quality

We use cluster cardinality to see number of samples in each cluster
```{r Clustering Cardinality}
plot_clust_cardinality <- cbind.data.frame(clusters_1, clusters_2) # Join clusters with  k =4 and k=6
names(plot_clust_cardinality) <- c("k_4", "k_5") # Set names
# Create bar plots
g_2 <- ggplot(plot_clust_cardinality, aes( x = factor(k_4))) + # Set x as cluster values
  geom_bar(stat = "count", fill = "steelblue") + # Use geom_bar with stat = "count" to count observations
    labs(x = "Cluster Number", y="Points in Cluster", # Set labels
         title = "Cluster Cardinality (k = 4)") +
  theme(panel.grid.major = element_blank(), # Turn of the background grid
        panel.grid.minor = element_blank(),
        panel.border = element_blank(),
        panel.background = element_blank()) 


g_3 <- ggplot(plot_clust_cardinality, aes( x = factor(k_5))) + # Set x as cluster values
  geom_bar(stat = "count", fill = "steelblue") + # Use geom_bar with stat = "count" to count observations
    labs(x = "Cluster Number", y="Points in Cluster", # Set labels
         title = "Cluster Cardinality (k = 5)") +
  theme(panel.grid.major = element_blank(), # Turn of the background grid
        panel.grid.minor = element_blank(),
        panel.border = element_blank(),
        panel.background = element_blank()) 
g_2
g_3
```

When k =4, only cluster 4 has significant amount of samples compared to other three clusters, which worths asking questions to analyze further. However, 5 clusters (k=5) did not help to solve this issue either. 

Next we will check on cluster magnitude by visualizing the within cluster sum of squares for each cluster to see how good a fit each of the clusters was fit. 


```{r Check Cluster Magnitude}
k_4_mag <- cbind.data.frame(c(1:4), fit_1$withinss) # Extract within cluster sum of squares
names(k_4_mag) <- c("cluster", "withinss") # Fix names for plot data
g_4 <- ggplot(k_4_mag, aes(x = cluster, y = withinss)) + # Set x as cluster, y as withinss
  geom_bar(stat = "identity", fill = "steelblue") + # Use geom bar and stat = "identity" to plot values directly
   labs(x = "Cluster Number", y="Total Point to Centroid Distance", # Set labels
         title = "Cluster Magnitude (k = 4)") +
  theme(panel.grid.major = element_blank(), # Turn of the background grid
        panel.grid.minor = element_blank(),
        panel.border = element_blank(),
        panel.background = element_blank()) 
  
k_5_mag <- cbind.data.frame(c(1:5), fit_2$withinss) # Extract within cluster sum of squares
names(k_5_mag) <- c("cluster", "withinss") # Fix names for plot data
g_5 <- ggplot(k_5_mag, aes(x = cluster, y = withinss)) +  # Set x as cluster, y as withinss
  geom_bar(stat = "identity", fill = "steelblue") + # Use geom bar and stat = "identity" to plot values directly
   labs(x = "Cluster Number", y="Total Point to Centroid Distance", # Set labels
         title = "Cluster Magnitude (k = 5)") +
  theme(panel.grid.major = element_blank(), # Turn of the background grid
        panel.grid.minor = element_blank(),
        panel.border = element_blank(),
        panel.background = element_blank()) 


g_4
g_5
```


When we have 4 clusters (k=4), both cluster 3 and 4 have similarities on the cluster sum of squares. Cluster 1 has the smallest value among all, while cluster 3 value is a little higher than cluster 1. 

When we have 5 clusters (k=5), cluster 4 has the highest value among all, while both cluster 1 and 5 have similarities on the cluster sum of squares. However, cluster 3 has the lowest value. 

And we also see with the number of cluster increasing, the cluster sum of squares actually decreases, which is the expected results we want to see. 

We can also plot both magnitude and cardinality on the graphs to see the relationship
```{r Magnitude v Cardinality}
k_4_dat <- cbind.data.frame(table(clusters_1), k_4_mag[,2]) # Join magnitude and cardinality
names(k_4_dat) <- c("cluster", "cardinality", "magnitude") # Fix plot data names

g_6 <- ggplot(k_4_dat, aes(x = cardinality, y = magnitude, color = cluster)) + # Set aesthetics
  geom_point(alpha = 0.8, size  = 4) +  # Set geom point for scatter
 geom_smooth(aes(x = cardinality, y = magnitude), method = "lm",
              se = FALSE, inherit.aes = FALSE, alpha = 0.5) + # Set trend  line
  labs(x = "Cluster Cardinality", y="Total Point to Centroid Distance", # Set labels
         title = "Cluster Magnitude vs Cardinality \n(k = 4)") +
  theme(panel.grid.major = element_blank(), # Turn of the background grid
        panel.grid.minor = element_blank(),
        panel.border = element_blank(),
        panel.background = element_blank()) 


k_5_dat <- cbind.data.frame(table(clusters_2), k_5_mag[,2]) # Join magnitude and cardinality
names(k_5_dat) <- c("cluster", "cardinality", "magnitude") # Fix plot data names

g_7 <- ggplot(k_5_dat, aes(x = cardinality, y = magnitude, color = cluster)) + # Set aesthetics
  geom_point(alpha = 0.8, size = 4) +  # Set geom point for scatter
  geom_smooth(aes(x = cardinality, y = magnitude), method = "lm",
              se = FALSE, inherit.aes = FALSE, alpha = 0.5) + # Set trend  line
  labs(x = "Cluster Cardinality", y="Total Point to Centroid Distance", # Set labels
         title = "Cluster Magnitude vs Cardinality \n(k = 5)") +
  theme(panel.grid.major = element_blank(), # Turn of the background grid
        panel.grid.minor = element_blank(),
        panel.border = element_blank(),
        panel.background = element_blank()) 


g_6
g_7
```

It seems that five cluster actually shows us a better positive relationship between magnitude and cardinality compared to four cluster. And since cluster 3(k=5) is far away from the fitted line, *cluster 3* might be anomalous clusters among all.  

Let's look at the heat map for all six stats on five clusters
```{r}
# Create vector of clusters
cluster2 <- c(1: 5)
# Extract centers
center_df_2 <- data.frame(cluster2, centers_2)

# Reshape the data

center_reshape_2 <- gather(center_df_2, # set the data 
                         stats, # the col name for the select columns' name
                         values, # the col name for the select columns' value
                         hp:speed) # select columns 

g_heat_2 <- ggplot(data = center_reshape_2, aes(x = stats, y = cluster2, fill = values)) +
  scale_y_continuous(breaks = seq(1, 5, by = 1)) +
  geom_tile() +
  coord_equal() + 
  theme_set(theme_bw(base_size = 22) ) +
  scale_fill_gradient2(low = "blue", # Choose low color
                       mid = "white", # Choose mid color
                       high = "red", # Choose high color
                       midpoint =0, # Choose mid point
                       space = "Lab", 
                       na.value ="grey", # Choose NA value
                       guide = "colourbar", # Set color bar
                       aesthetics = "fill") + # Select aesthetics to apply
  coord_flip()
g_heat_2

```
It seems that cluster 3 is generally better than any other clusters and cluster 4 performs the lowest in terms of all the stats values. We can also subset our original pokemon by each cluster number and create five cluster subset. 

```{r}
group_1 <- pokemon_list[clusters_2==1,]  # subset cluster 1 
group_2 <- pokemon_list[clusters_2==2,]  # subset cluster 2 
group_3 <- pokemon_list[clusters_2==3,]  # subset cluster 3
group_4 <- pokemon_list[clusters_2==4,]  # subset cluster 4
group_5 <- pokemon_list[clusters_2==5,]  # subset cluster 5
```

```{r}
make_ridge_plot <- function(data){
  
  data <- data %>%
      pivot_longer(cols=hp:speed,   # transform into long table
                   names_to="stats_type",
                   values_to="score")
  
  ggplot(data,aes(x=score, y=stats_type,fill=stats_type)) +
    geom_boxplot() +
    scale_fill_viridis(discrete = TRUE, alpha=0.8, option="A") +
    geom_jitter(color="black", size=0.4, alpha=0.) +
    theme_ipsum() +
    theme(
      legend.position="none",
      plot.title = element_text(size=12)
    ) +
    ggtitle("All Pokemon in Galar Region Stats Boxplot") +
    xlab("Stats Score")+
    ylab("")
}

rp3 <- make_ridge_plot(group_3)
rp4 <- make_ridge_plot(group_4)
```

```{r,fig.width=10}
library(ggpubr)
ggarrange(rp3, rp4,
          ncol = 2, nrow = 1)
```


### Method 2 : Calculate Optimal Cluster Number (Gap Statistic Method)
```{r}
library(cluster)
# compute gap statistic
set.seed(111111)
gap_stat <- clusGap(pokemon_new[,2:7], FUN = kmeans, nstart = 25,
                    K.max = 10, B = 50)
# Print the result
print(gap_stat, method = "firstmax")

fviz_gap_stat(gap_stat)
```
If we use gap statistic method, the optimal cluster will be three only. We will not spend more time on this method at this time. 

## Hierarchical Clustering


```{r}
library(cluster)    # clustering algorithms
library(factoextra) # clustering visualization
library(dendextend) # for comparing two dendrograms
library(circlize) # For circular dendrograms
```

```{r,fig.width=8}
# set pokemon name as index
rownames(pokemon_new) <- pokemon_new$pokemon_name

# Calculate distances between points
dist_mat <- dist(pokemon_new[,2:7], # Set dataset
                 method = "euclidean") # Set distance measure to use

# Run hierarchical clustering
hc <- hclust(dist_mat, # Set distance matrix to use 
              method = "average" ) # Set linkage measure to use

plot(hc, # Set hierarchical clustering as plot object
     cex = 0.6, # Set text size
     hang = -1 ) # Set label position
```

```{r,fig.width=8}
plot(hc, # Set hierarchical clustering as plot object
     cex = 0.6,  # Set text size
     hang = -1) # Set label position
# Draw rectangles around branches/clusters
rect.hclust(hc, # Set hierarchical clustering object
            k = 8, # Set number of clusters 
            border = 2:6) # Sets colors for rectangles
```

```{r,fig.width=8}
# Create dendrogram
dend <- as.dendrogram(hc)
# order it the closest we can to the order of the observations:
dend <- rotate(dend, 1:50)

# Color the branches based on the clusters:
dend <- color_branches(dend, k=8) 

# We hang the dendrogram a bit:
dend <- hang.dendrogram(dend,hang_height=0.1)
# reduce the size of the labels:
dend <- set(dend, "labels_cex", 0.55)
# And plot:
par(mar = c(3,3,3,7))
plot(dend, 
     main = "Clustered Pokemon Data",
     horiz =  TRUE,  nodePar = list(cex = .007))
```

```{r,fig.height=10}
par(mar = rep(0,4))
circlize_dendrogram(dend)
```

## Factor Analysis

```{r}
# prepare dataset for factor analysis
library(psych)
pokemon_factor_id <- pokemon_list[,c(1,2,7:12)] # extract six stats with id and name
pokemon_factor <- pokemon_list[,7:12] # extract only six stats for factor analysis
```


```{r}
# Determine Number of Factors to Extract
library(nFactors)
ev <- eigen(cor(pokemon_factor)) # get eigenvalues
ap <- parallel(subject=nrow(pokemon_factor),var=ncol(pokemon_factor),
  rep=100,cent=.05)
nS <- nScree(x=ev$values, aparallel=ap$eigen$qevpea)
plotnScree(nS) 
```
Since two methods suggest choosing three factors and the other two suggest choosing only one factor. Let's try number of factors from 3 to 1, and see how they perform. 

```{r}
fa(pokemon_factor, 3) # three factors
fa(pokemon_factor, 2) # two factors
fa(pokemon_factor, 1) # one factor
model <- fa(pokemon_factor, 1)
```
It seems that one factor MR1 is sufficient enough to capture correlation coefficient for six variables.

```{r}
pokemon_scores <- cbind(model$scores,pokemon_factor_id)

colnames(pokemon_scores)
pokemon_scores$total <- rowSums(pokemon_scores[,4:9], na.rm=TRUE)

pokemon_scores <- pokemon_scores[order(-pokemon_scores$total),] 
```

# ```{r}
# # Maximum Likelihood Factor Analysis
# # entering raw data and extracting 3 factors,
# # with varimax rotation
# fit <- factanal(pokemon_factor, 3, rotation="varimax")
# print(fit, digits=2, cutoff=.3, sort=TRUE)
# # plot factor 1 by factor 2
# load <- fit$loadings[,1:2]
# plot(load,type="n") # set up plot
# text(load,labels=names(pokemon_factor),cex=.7) # add variable names 
# ```

We can see that the attack variable loads heavily close to 1 on Factor 1 but has low loading on Factor 2. We see that the defense variable loads heavily around 0.8 on Factor 2  but not on Factor 1. Thus, we can define the factors using those terms. Factor 1 might be labeled attack, and Factor 2 might be labeled defense.

# ```{r}
# # Pricipal Components Analysis
# # entering raw data and extracting PCs
# # from the correlation matrix
# fit <- princomp(pokemon_factor, cor=TRUE)
# summary(fit) # print variance accounted for
# loadings(fit) # pc loadings
# plot(fit,type="lines") # scree plot
# fit$scores # the principal components
# biplot(fit) 
# ```

